Powdery mildew yield loss mitigation strategies in mungbean, a meta-analysis.

### 2020-06-15 03_Prep_data.Rmd
add lme to test dose effect in models

```{r load_libraries, include=FALSE}
if (!require("pacman"))
   install.packages("pacman")
pacman::p_load(tidyverse, kableExtra, agricolae
)
```

@@

```{r variance_types}
PM_MB_means %>%
   filter(fungicide_ai == "tebuconazole" |
             fungicide_ai == "propiconazole") %>%
   group_by(trial_ref, location, Y_error_type, fungicide_ai) %>%
   summarise(DMI_trts = length(Y_error_type)) %>%
   arrange(Y_error_type) %>%
   kable(caption = "Number of treatments for each trial and the error type reported") %>%
   #kable_styling(fixed_thread = T) %>% 
   footnote(general = "DMI - demethylation inhibitors (tebuconazole and propiconazole).")
```

@@

First lets get a list of trials that reported LSD.  
```{r all_LSD_trials}
Trials <- PM_MB_means %>%
   filter(Y_error_type == "lsd (P=0.05)") %>%
   select(trial_ref) %>%
   distinct()
```

@@

```{r cluster_clusters}
clusters <-
   unique(slim_PM_dat[c("cluster_1", "cluster_2", "cluster_3")])

@@

Single sprays either early, recommended timing or late.

Two or more sprays where the first spray was either early, recommended timing, or late.

@@

```{r tableSprayManagement}
table(slim_PM_dat$spray_management)

@@

There are few 'Early_plus' and 'late' treatments, these treatments therefore will have too few comparisons with other treatments in the meta-analysis to be provide accurate results. Therefore we will remove 'Early_plus' and 'Late' from the analysis.

```{r simpler_clusters_REMOVE}
slimmer_PM_dat <-
   slim_PM_dat[#slim_PM_dat$spray_management != "Late" &
                  slim_PM_dat$spray_management != "Early_plus", ]

write.csv(slimmer_PM_dat, file = "cache/slimmer_PM_clusterdat.csv", row.names = FALSE)
```

@@

The linear model confirms a significant positive relationship between grain yield and AUDPC. 
This is not what we would expect and the low R squared value of `r round(summary(lm(grain_yield.t.ha. ~ AUDPC_m, data = PM_MB_means))$adj.r.squared,3)` perhaps there are additional factors influencing this relationship.

@@

```{r}

@@

>>>>

```{r load_libraries, include=FALSE}
if (!require("pacman"))
   install.packages("pacman")
pacman::p_load(tidyverse, kableExtra, agricolae, lme4
)
```

@@

```{r variance_types}
PM_MB_means %>%
   filter(fungicide_ai == "tebuconazole" |
             fungicide_ai == "propiconazole") %>%
   group_by(trial_ref, location, Y_error_type, fungicide_ai) %>%
   summarise(DMI_treatments = length(Y_error_type)) %>%
   arrange(Y_error_type) %>%
   select(!fungicide_ai)%>%
   kable(caption = "Number of treatments for each trial and the error type reported", align = "c") %>%
   footnote(general = "DMI - demethylation inhibitors (tebuconazole and propiconazole).")
```

@@

First lets get a list the of trials that reported LSD.  
```{r all_LSD_trials}
Trials <- PM_MB_means %>%
   filter(Y_error_type == "lsd (P=0.05)") %>%
   select(trial_ref) %>%
   distinct()
```

@@

```{r cluster_clusters, warning=FALSE}
clusters <-
   unique(slim_PM_dat[c("cluster_1", "cluster_2", "cluster_3")])

@@

Single sprays of either:  
   a) early,  
   b) recommended timing,  
   c) late.

Two or more sprays where the first spray was either:  
   d) early,  
   e) recommended timing,  
   f) late.

@@

```{r tableSprayManagement}
table(slim_PM_dat$spray_management)
```

There are few 'Early_plus' treatments, these treatments therefore will have too few comparisons with other treatments in the meta-analysis to provide accurate results. Therefore we will remove 'Early_plus' from the analysis.

```{r simpler_clusters_REMOVE}
slimmer_PM_dat <-
   slim_PM_dat[slim_PM_dat$spray_management != "Early_plus", ]

@@

## Test dose effect

Earlier we noted there were different doses of propiconazole which might influence our analysis. 
Lets check this by comparing a linear mixed effect model including and excluding dose. 
We already have a factor breaking down the dose in the `trial` variable, lets make a new one without dose.

```{r dose_test}
slimmer_PM_dat$trial_noDose <- paste(slimmer_PM_dat$trial_ref, slimmer_PM_dat$year,
                 slimmer_PM_dat$location,slimmer_PM_dat$host_genotype,
                 slimmer_PM_dat$row_spacing,sep = "_")

cbind(Dose = head(slimmer_PM_dat$trial),
      NoDose = head(slimmer_PM_dat$trial_noDose)) %>%
   kable()
```

Lets test the dose effect in a basic linear mixed effect model

```{r dose_lmer}

m1 <-
   lmer(log(grain_yield.t.ha * 1000) ~ factor(spray_management) +
           (1 | trial),
        data = slimmer_PM_dat)

m2 <-
   lmer(log(grain_yield.t.ha * 1000) ~ factor(spray_management) +
           (1 | trial_noDose),
        data = slimmer_PM_dat)

anova(m1,m2)

Using the model without dose produces a better fit, as shown by the lower AIC and is significantly different from the model `m1`. 
We will remove the dose term from the trial grouping.

```{r}
slimmer_PM_dat$trial <- slimmer_PM_dat$trial_noDose
slimmer_PM_dat <- slimmer_PM_dat[,colnames(slimmer_PM_dat) != "trial_noDose"] #remove second trial column
write.csv(slimmer_PM_dat, file = "cache/slimmer_PM_clusterdat.csv", row.names = FALSE)
```

@@

This is not what we would expect and the low R squared value of `r round(summary(lm(grain_yield.t.ha. ~ AUDPC_m, data = PM_MB_means))$adj.r.squared,3)` perhaps there are additional factors influencing this relationship.

@@

```{r}
slimmer_PM_dat <- read.csv(here("cache/slimmer_PM_clusterdat.csv"))

@@

write.csv(slimmer_PM_dat, file = here("cache/slimmer_PM_clusterdat.csv"), row.names = FALSE)
```
